---
title: "Midterm Assignment"
author: "Jiaqi Sun"
date: "2022-11-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(magrittr)
library(readxl)
library(dplyr)
```

#Tidying the data
```{r}
## Start by reading the data
strawb <- read_xlsx("strawberries-2022oct30-a.xlsx",col_names = TRUE)

## Get the column names and index them
cnames <- colnames(strawb)
x <- 1:dim(strawb)[2]

## Column 1 contains two unique values.  
## Retain column 1 -- those values might be needed.
unique(strawb[1])

## Column 2 -- contains the years included in this dataset.
## Keep column 2, of course.
unique(strawb[2])

## Column 3 -- contains the time periods covered by in the dataset.
## There's only one -- years.  No info here.  Drop it
unique(strawb[3])

## Set T as an indicator
T <- NULL

## Collect number of unique rows in each column
for(i in x){T <- c(T, dim(unique(strawb[i]))[1])}

## Use T to select columns to drop -- 
drop_cols <- cnames[which(T == 1)]

## Now, drop the columns with only one unique value.
strawb %<>% select(!all_of(drop_cols))

## Let's arrange the data frame by year and state.
strawb %<>% arrange(Year, State)
```

#Tidying the data
```{r}
strawb %<>% separate(col=`Data Item`,
                    into = c("Strawberries", "type", "items", "units"),
                    sep = ",",
                    fill = "right")
```

#Tibble Organic
```{r}
type_organic <- grep("organic", 
               strawb$type, 
               ignore.case = T)

Domain_organic <- grep("organic", 
                      strawb$Domain, 
                      ignore.case = T)

org_rows <- intersect(type_organic, Domain_organic)

strawb_organic <- strawb %>% slice(org_rows, preserve = FALSE)

strawb_organic[strawb_organic$type == " ORGANIC - SALES", ]$units = strawb_organic[strawb_organic$type == " ORGANIC - SALES", ]$items


```

#Tibble Non-Organic
```{r}
strawb_non_organic <- strawb %>% filter(!row_number() %in% org_rows)

strawb_non_organic[strawb_non_organic$Strawberries == "STRAWBERRIES - PRICE RECEIVED", ]$items = strawb_non_organic[strawb_non_organic$Strawberries == "STRAWBERRIES - PRICE RECEIVED", ]$type
```

#Tibble Chemical
```{r}
## the chemical data in the column "type" 
## is in the rows where type = BEARING - Applications

chem_rows <- grep("BEARING - APPLICATIONS", 
                       strawb_non_organic$type, 
                       ignore.case = T)

strawb_chem <- strawb_non_organic %>% slice(chem_rows, preserve = FALSE)
```

#Tidying the Tibble Chemical
```{r}
## now clean up the workspace before tackling the three tibbles just created.
rm(x, T, drop_cols, type_organic, Domain_organic, org_rows, chem_rows,
  cnames, i)

  before_cols = colnames(strawb_chem)
  T = NULL
  x = length(before_cols)

   for(i in 1:x){
    b <- length(unlist(strawb_chem[,i] %>% unique()) )
    T <- c(T,b)
   }
    
  drop_cols <- before_cols[which(T == 1)]
  strawb_chem %<>% select(!all_of(drop_cols))
  after_cols = colnames(strawb_chem)

## drop_cols is labeling information
  
  temp1 <- strawb_chem %>% select(units) %>% distinct()
  ## in units rows are either NA or AVG
  
## separate Domain Category at :
  
  strawb_chem %<>% separate(col=`Domain Category`, 
                                    into = c("dc1", "chem_name"),
                                    sep = ":", 
                                    fill = "right")
  
temp1 <- strawb_chem %>% select(chem_name) %>% unique()
length(unlist(temp1))

## first clean it up

aa  <- grep("measured in", 
                    strawb_chem$items, 
                    ignore.case = T)
length(aa)

## so "item" is really units

## are Domain and dc1 the same thing?

sum(strawb_chem$Domain == strawb_chem$dc1) == dim(strawb_chem)[1]

## yes, they are. let's work with dc1. Also, drop State ANSI

strawb_chem %<>% select(Year, State, items, units, dc1, chem_name, Value)

## items should only contain unit names
## rename units to category

strawb_chem %<>% rename(category = units)

## remove "MEASURED IN " 

strawb_chem$items <- str_remove_all(strawb_chem$items, "MEASURED IN ")

## rename items to units
strawb_chem %<>% rename(units = items)

## Do all the dc1 entries begen with "Chemical"?

bb  <- grep("CHEMICAL, ", 
            strawb_chem$dc1, 
            ignore.case = T)
length(bb)
chem <- 1:2112

non_chem_rows <- setdiff(chem, bb)
length(non_chem_rows)

## on let's look at these rows in a tibble

temp1 <- strawb_chem %>% slice(non_chem_rows)

### !! fertilizers  

## keep them -- probably won't use them as alone tibble

fertilizers <- temp1

## now remove "CHEMICAL, " from the entries in the dc1
## and rename the column chem_types

strawb_chem$dc1 <- str_remove_all(strawb_chem$dc1, "CHEMICAL, ")

strawb_chem$dc1 %>% unique()

strawb_chem %<>% rename(chem_types = dc1)

## now fix the chem_name column

## remove the parens

strawb_chem$chem_name <- str_remove_all(strawb_chem$chem_name, "\\(")

strawb_chem$chem_name <- str_remove_all(strawb_chem$chem_name, "\\)")

## separate chem_name and chem_code

strawb_chem %<>% separate(col = chem_name,
                          into = c("chem_name","chem_code"),
                          sep = "=",
                          fill = "right"
                            ) 

## now fill in a label fot NA in the category column
aa <- which(strawb_chem$units == " LB")

bb <- which(is.na(strawb_chem$category))

sum(aa==bb)==length(aa)

strawb_chem$category[bb] = "Total"

rm(aa, after_cols, b, bb, before_cols, drop_cols, i, non_chem_rows, T, x)

```

#More tidy work to generate columns as "Per Acre", "Average per Acre per Application" and "Average per Acre per Year"
```{r}
strawb_chem %>% 
  select(-chem_code) %>% 
  pivot_wider(names_from = units, values_from = Value) %>% 
  rename("per Acre" = " LB") %>% 
  rename("Average per Acre per Application" =  " LB / ACRE / APPLICATION") %>% 
  rename("Average per Acre per Year" =  " LB / ACRE / YEAR") -> strawb_chem
```

#understand the units used in the data set 
```{r}
unique(strawb_chem$category)

#we measure unites in 3 ways: 
#Total(per acre), average per acre per year, average per acre per application

#A hundredweight (abbreviated as CWT) is a standard unit of weight or mass used in certain commodities markets. In North America, a hundredweight is equal to 100 pounds(LB) 

#We see that in tibble "strawb_organic", the units are " MEASURED IN $" and " MEASURED IN CWT", while in the tibble"strawb_non_organic", the units are showed as " MEASURED IN $ / CWT", we have to measure in the same unites for better comparison
```

#Chemical Investigation
## investigate chemicals commonly used according to "Shopper’s Guide to Pesticides in Produce™" wihch are deadly possions
```{r}
chemical_num <- length(unique(strawb_chem$chem_name))

#Bifenthrin, 27 results found, possible human carcinogen
bb  <- grep("BIFENTHRIN", 
            strawb_chem$chem_name, 
            ignore.case = T)

bifen <- strawb_chem %>% slice(bb)

#Carbendazim, nothing found，ormone-disrupting fungicide that damages the male reproductive system and that the EU has banned because of safety concerns.
cc <- grep("CARBENDAZIM", 
            strawb_chem$chem_name, 
            ignore.case = T) 

#chloropicrin, 18 results found，The EU has banned it
ch  <- grep("chloropicrin", 
            strawb_chem$chem_name, 
            ignore.case = T)

chloropicrin <- strawb_chem %>% slice(ch)

#methyl bromide, 2 results found， The most notorious strawberry fumigant
me  <- grep("methyl bromide", 
            strawb_chem$chem_name, 
            ignore.case = T)

methyl <- strawb_chem %>% slice(me)

#1,3-dichloropropene, nothing found， The EU has banned it
di  <- grep("1,3-dichloropropene", 
            strawb_chem$chem_name, 
            ignore.case = T)

#Telone, nothing found, 1,3-dichloropropene, The EU has banned it
te  <- grep("Telone", 
            strawb_chem$chem_name, 
            ignore.case = T)
```

#Poissons used per acre? per year? per application? 
##Do these quantities vary by state? by year?
```{r}
#BIFENTHRIN used per acre per application, varied by year and state
bifen %>% 
  filter(`per Acre` > 0) %>% 
  select(Year, State, chem_name, `per Acre`) %>% 
  group_by(Year, State) -> bifen_pa

#BIFENTHRIN used per acre per application, varied by year and state
bifen %>% 
  filter(`Average per Acre per Application` > 0) %>% 
  select(Year, State, chem_name, `Average per Acre per Application`) %>% 
  group_by(Year, State) -> bifen_papa

#BIFENTHRIN used per acre per year
bifen %>% 
  filter(`Average per Acre per Year` > 0) %>% 
  select(Year, State, chem_name, `Average per Acre per Year`) %>% 
  group_by(Year, State) ->bifen_papy

#chloropicrin used per acre, varied by year and state
chloropicrin %>% 
  filter(`per Acre` > 0) %>% 
  select(Year, State, chem_name, `per Acre`) %>% 
  group_by(Year, State) -> chloropicrin_pa

#chloropicrin used per acre per application, varied by year and state
chloropicrin %>% 
  filter(`Average per Acre per Application` > 0) %>% 
  select(Year, State, chem_name, `Average per Acre per Application`) %>% 
  group_by(Year, State) -> chloropicrin_papa

#chloropicrin used per acre per year, varied by year and state
chloropicrin %>% 
  filter(`Average per Acre per Year` > 0) %>% 
  select(Year, State, chem_name,`Average per Acre per Year`) %>% 
  group_by(Year, State) -> chloropicrin_papy

#methyl used per acre per application, varied by year and state
#nothing returned (with footnote)

pa <- rbind(bifen_pa, chloropicrin_pa)
papa <- rbind(bifen_papa, chloropicrin_papa)
papy <- rbind(bifen_papy, chloropicrin_papy)

pois_all <- rbind(bifen_pa,bifen_papa,bifen_papy, chloropicrin_pa, chloropicrin_papa, chloropicrin_papy)

```


## investigate chemicals commonly used，wihch are safe ones
```{r}

#Acibenzolar-S-methyl, 12 found, has little toxicity to mammals with an LD50 of over 2000 mg/kg (rats, oral). However, it can cause moderate eye irritation.
acib  <- grep(" ACIBENZOLAR-S-METHYL ", 
            strawb_chem$chem_name, 
            ignore.case = T)

ACIB <- strawb_chem %>% slice(acib)

#Acetamiprid, 16 found, has been classified by the EPA as “unlikely” to be a human carcinogen.
acet  <- grep(" ACETAMIPRID ", 
            strawb_chem$chem_name, 
            ignore.case = T)

ACET <- strawb_chem %>% slice(acet)

#acequinocyl, 10 found, unlikely to present a risk to consumer health.
aceq  <- grep(" ACEQUINOCYL ", 
            strawb_chem$chem_name, 
            ignore.case = T)

ACEQ <- strawb_chem %>% slice(aceq)
```

#Building tibbles given graph"3-Distribution-of-Strawberry-Production-in-the-US-by-state-Bertelsen-1995.png"
##Tibble California
```{r}
strawb_chem_ca <- strawb_chem %>% 
  filter(State == "CALIFORNIA")
```
##Tibble FLORIDA
```{r}
strawb_chem_fl <- strawb_chem %>% 
  filter(State == "FLORIDA")
```
##Tibble OREGON
```{r}
strawb_chem_or <- strawb_chem %>% 
  filter(State == "OREGON")
```

```{r}
# plot
# load library
library(ggplot2)

# Create test data.
hazardous_chems <- data.frame(
  category=c("Bifenthrin", "chloropicrin", "methyl bromide"),
  count=c(27, 18, 3)
)
 
# Compute percentages
hazardous_chems$fraction <- hazardous_chems$count / sum(hazardous_chems$count)

# Compute the cumulative percentages (top of each rectangle)
hazardous_chems$ymax <- cumsum(hazardous_chems$fraction)

# Compute the bottom of each rectangle
hazardous_chems$ymin <- c(0, head(hazardous_chems$ymax, n=-1))

# Compute label position
hazardous_chems$labelPosition <- (hazardous_chems$ymax + hazardous_chems$ymin) / 2

# Compute a good label
hazardous_chems$label <- paste0(hazardous_chems$category, "\n value: ", hazardous_chems$count)

# Make the plot
ggplot(hazardous_chems, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
  geom_rect() +
  geom_label( x=3.5, aes(y=labelPosition, label=label), size=4) +
  scale_fill_brewer(palette=4) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void() +
  theme(legend.position = "none")
```

#Chemical graphic exploration

```{r}
#Florida uses far more less poisson chemicals than California and it stopped using poisson chemicals gradually in recent years

#general use of chemicals
ggplot(pois_all, aes(x = chem_name, fill = State)) + geom_bar() + facet_wrap(~Year)

#Average per Acre per Application
ggplot(papa, aes(x = as.double(`Average per Acre per Application`), fill = State)) + geom_histogram(binwidth = 180) +
  facet_wrap(~Year)

#Average per Acre per Year
ggplot(papy, aes(x = as.double(`Average per Acre per Year`), fill = State)) + geom_histogram(binwidth = 180) +
  facet_wrap(~Year)

#Average per Acre 
ggplot(pa, aes(x = as.double(`per Acre`), fill = State)) + geom_histogram(binwidth = 1000000) +
  facet_wrap(~Year)
```

